---
title: "Log2_Heatmap"
output: html_document
date: "2025-05-14"
---

#Normalized ASV Table Preparation
```{r , eval=FALSE}
rm(list = ls())
gc()
suppressMessages({
  library("tidyverse")
  library("phyloseq")
  library("rstatix")
  library("vegan")
  library("picante")
  library("ggpubr")
  library(vegan)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(microbiomeutilities)
  library(viridis)
  library(patchwork)
  library(pheatmap)
  library(RColorBrewer)
  library(reshape2)
})
no_of_cores = 16
setwd("/work/benson/bpeng4/Givaudan/Givaudan_All_16S")
load("intermediate/Givaudian_ps.rda")
ls()
ps

#Remove Fs
ps <- subset_samples(ps , Category != "FS")
#Remove Spaces
# Remove leading/trailing whitespace from all levels in the Category column
ps@sam_data$Category <- trimws(ps@sam_data$Category)

###Pre-arrangement on  the phyloseq file
################
ps.clean <- subset_taxa(ps, Kingdom == "Bacteria") %>%
  subset_taxa(!is.na(Phylum)) %>%
  subset_taxa(!Class %in% c("Chloroplast")) %>%
  subset_taxa(!Family %in% c("Mitochondria"))
ps.clean
#ps.clean.re <- transform_sample_counts(ps.clean, function(x) x / sum(x))
#Filter Out Taxa Exist in At Least 25% samples
ps.clean.p0 <- filter_taxa(ps.clean, function (x) {sum(x > 0) >= 251}, prune=TRUE)
ps.clean.p0

ps.rare<-rarefy_even_depth(ps.clean.p0, sample.size = 6138,
                           rngseed = 111, replace = TRUE, trimOTUs = TRUE, verbose = TRUE)

ps.rare<-subset_samples(ps.rare, !ps.rare@sam_data$Category == "FBB16")

ps.genus = phyloseq::tax_glom(ps.rare, taxrank = rank_names(ps.rare)[6])
```

#Write a Function for Permanova Test Filteration
```{r}
ps.filter<-function(ps.genus){
  #Extract OTU/ASV table and metadata:
  otu_table <- as.data.frame(otu_table(ps.genus))
  metadata <- data.frame(sample_data(ps.genus))  # Force conversion
  taxa_table<- as.data.frame(tax_table(ps.genus))

  #Use the adonis function (PERMANOVA) from vegan to analyze treatment(Sample) effects
  metadata$Sample <- as.factor(metadata$Sample)  # Ensure it's a factor
  adonis_results <- adonis2(otu_table ~ Sample, data = metadata, permutations = 999, method = "bray")
  
  #Extract p-values
  p_value <- adonis_results$`Pr(>F)`[1]  # Extract p-value for treatment
  
  #test each taxon individually:
  p_values <- apply(otu_table, 2, function(x) {
    df <- data.frame(x = x, Sample = metadata$Sample)
    fit <- aov(x ~ Sample, data = df)
    summary(fit)[[1]][["Pr(>F)"]][1]  # Extract p-value
  })
  
  #Adjust for multiple testing (FDR correction):
  p_values_adj <- p.adjust(p_values, method = "fdr")
  
  #Mark taxa as significant if p < 0.05
  signif_taxa <- names(p_values_adj[p_values_adj < 0.05])

  #Filter for only the significant taxa
  ps.genus.filtered <- prune_taxa(taxa_names(ps.genus) %in% signif_taxa, ps.genus)
  
  return(ps.genus.filtered)
}

```

#Write a Function for plotting log2 fold Heatmap for Category
```{r, eval=FALSE}
heatmapcategory <- function(ps.genus) {
  library(phyloseq)
  library(dplyr)
  library(tibble)
  library(RColorBrewer)
  
  # Extract OTU (abundance) table and metadata from phyloseq
  otu_table_df <- as.data.frame(otu_table(ps.genus))
  meta_df <- data.frame(sample_data(ps.genus))
  
  # Add sample and category info to OTU table
  otu_table_df$Sample <- meta_df$Sample
  otu_table_df$Category <- meta_df$Category
  
  # Aggregate by sample and category (mean abundance per sample)
  otu_table_avg <- otu_table_df %>%
    group_by(Category, Sample) %>%
    summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)), .groups = "drop")
  
  # Set Sample as rownames before dropping it from columns
  otu_table_avg <- otu_table_avg %>%
    column_to_rownames(var = "Sample")
  
  # Remove 'Category' column, then convert to matrix
  otu_table_matrix <- otu_table_avg[, !colnames(otu_table_avg) %in% "Category"] %>%
    as.matrix()
  
  # Create a phyloseq OTU table object
  otu_table_avg_phylo <- otu_table(otu_table_matrix, taxa_are_rows = FALSE)
  
  # Prepare corresponding sample metadata
  meta_avg_phylo <- meta_df %>%
    distinct(Sample, .keep_all = TRUE)
  rownames(meta_avg_phylo) <- meta_avg_phylo$Sample
  meta_avg_phylo <- sample_data(meta_avg_phylo)
  
  # Match taxa between original and averaged OTU tables
  common_taxa <- intersect(rownames(tax_table(ps.genus)), colnames(otu_table_matrix))
  tax_table_filtered <- tax_table(ps.genus)[common_taxa, ]
  
  # Create a new phyloseq object
  ps.genus.avg <- phyloseq(otu_table_avg_phylo, tax_table_filtered, meta_avg_phylo)
  
  # Step 2: Apply log2(x + 1)
  ps.genus.log2 <- transform_sample_counts(ps.genus.avg, function(x) log2(x + 1))

  # Step 3: Convert log2(x + 1) to 10^(log2(x + 1))
  ps.genus.log2.10 <- transform_sample_counts(ps.genus.log2, function(x) 10^x)
  
  # Plot heatmap of top 50 taxa
  heat.sample <- plot_taxa_heatmap(
    ps.genus.log2.10,
    subset.top = 50,
    VariableA = c("Category"),
    heatcolors = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(100),
    transformation = "log10"
  )
  
  return(heat.sample)
}


```

#Plot
```{r}
ps.genus.filtered <- ps.filter(ps.genus)

heatmapcategory(ps.genus.filtered)

```


#Write a Function for plotting log2 fold Heatmap for Sample
```{r}
heatmapsample<-function(ps.genus){
  # Extract OTU (abundance) table and metadata from phyloseq
  otu_table_df <- as.data.frame(otu_table(ps.genus))
  meta_df <- data.frame(sample_data(ps.genus))
  
  # Add category information to the OTU table
  otu_table_df$Sample <- meta_df$Sample
  otu_table_df$Category <- meta_df$Category
  
  # Aggregate by sample (mean abundance per sample)
  otu_table_avg <- otu_table_df %>%
    group_by(Category, Sample) %>%
    summarise(across(where(is.numeric), mean, na.rm = TRUE))
  
  # Convert back to matrix and ensure row names are correct
  otu_table_avg <- otu_table_avg[, !colnames(otu_table_avg) %in% "Category"]
  otu_table_avg <- column_to_rownames(otu_table_avg, var = "Sample") %>%
    as.matrix()
  
  otu_table_avg_phylo <- otu_table(otu_table_avg, taxa_are_rows = FALSE) 
  
  
  # Create a new sample metadata table
  meta_avg_phylo <- meta_df %>% distinct(Sample, .keep_all = TRUE)
  meta_avg_phylo$Category <- meta_avg_phylo$Category
  rownames(meta_avg_phylo) <- meta_avg_phylo$Sample
  meta_avg_phylo <- sample_data(meta_avg_phylo)
  
  # Ensure tax_table matches the taxa in otu_table_avg_phylo
  common_taxa <- intersect(rownames(tax_table(ps.genus)), colnames(otu_table_avg_phylo))
  tax_table_filtered <- tax_table(ps.genus)[common_taxa, ] 
  
  # Create a new phyloseq object
  ps.genus.avg <- phyloseq(otu_table_avg_phylo, tax_table_filtered, meta_avg_phylo)
  
  # Step 2: Apply log2(x + 1)
  ps.genus.log2 <- transform_sample_counts(ps.genus.avg, function(x) log2(x + 1))

  # Step 3: Convert log2(x + 1) to 10^(log2(x + 1))
  ps.genus.log2.10 <- transform_sample_counts(ps.genus.log2, function(x) 10^x)
  
  # Plot heatmap of top 50 taxa
  heat.sample <- plot_taxa_heatmap(
    ps.genus.log2.10,
    subset.top = 50,
    VariableA = c("Sample"),
    heatcolors = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(100),
    transformation = "log10"
  )
  }
```

#Plot
```{r}
heatmapsample(ps.genus.filtered)
```

heatmapsample(ps.genus.filtered)


#Ouput the Corresponding(log2 ratio) OTU table
```{r}
  # Extract OTU (abundance) table and metadata from phyloseq
  otu_table_df <- as.data.frame(otu_table(ps.genus.filtered))
  meta_df <- data.frame(sample_data(ps.genus.filtered))
  
  # Add category information to the OTU table
  otu_table_df$Sample <- meta_df$Sample
  otu_table_df$Category <- meta_df$Category
  
  # Aggregate by sample (mean abundance per sample)
  otu_table_avg <- otu_table_df %>%
    group_by(Category, Sample) %>%
    summarise(across(where(is.numeric), mean, na.rm = TRUE))
  
  # Convert back to matrix and ensure row names are correct
  otu_table_avg <- otu_table_avg[, !colnames(otu_table_avg) %in% "Category"]
  otu_table_avg <- column_to_rownames(otu_table_avg, var = "Sample") %>%
    as.matrix()
  
  otu_table_avg_phylo <- otu_table(otu_table_avg, taxa_are_rows = FALSE) 
  
  
  # Create a new sample metadata table
  meta_avg_phylo <- meta_df %>% distinct(Sample, .keep_all = TRUE)
  meta_avg_phylo$Category <- meta_avg_phylo$Category
  rownames(meta_avg_phylo) <- meta_avg_phylo$Sample
  meta_avg_phylo <- sample_data(meta_avg_phylo)
  
  # Ensure tax_table matches the taxa in otu_table_avg_phylo
  common_taxa <- intersect(rownames(tax_table(ps.genus)), colnames(otu_table_avg_phylo))
  tax_table_filtered <- tax_table(ps.genus)[common_taxa, ] 
  
  # Create a new phyloseq object
  ps.genus.avg <- phyloseq(otu_table_avg_phylo, tax_table_filtered, meta_avg_phylo)
  
  # Step 2: Apply log2(x + 1)
  ps.genus.log2 <- transform_sample_counts(ps.genus.avg, function(x) log2(x + 1))
  
  #Write a function to Output the OTU table Corresponding to the Heatmap
  OutOTU<-function(ps.run, filename = "otu_otuput"){
  OTU<-otu_table(ps.run) |>as.data.frame()
  Taxa<-tax_table(ps.run) |>as.data.frame()
  Taxa$Name<- apply(Taxa[,1:6], 1, function(x) paste(x, collapse = "_"))
  colnames(OTU)<-Taxa$Name
  OTU$Total<- rowSums(OTU)
  return(OTU)
}

OTU<-OutOTU(ps.genus.log2)
```

write.csv(OTU, file = paste0("/work/benson/bpeng4/Givaudan/Plot/otu.sample.log2.csv"))
